---
title: "cross_validation"
author: "Jiayi Shi"
date: "2022-11-15"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

## Model selection
nested models: anova()
non-nested models: AIC/BIC/etc
balance bwt goodness of fit and complexity

## CV by hand
build model on training dataset and evaluate using RMSE on test dataset

Generates data under a non-linear model

```{r}
nonlin_df = 
  tibble(
    id = 1:100,
    x = runif(100, 0, 1),
    y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
  )

nonlin_df %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point()
```

Split into training and test sets using:

df %>% `sample_n(numebr_in_train)`/`slice_sample(n = number_in_train)`: select number_in_train observations from the original dataset `df` randomly.

and

`anti_join(x,y)`: return all rows from x without a match in y.

```{r}
train_df = nonlin_df %>% slice_sample(prop = 0.8)
test_df = anti_join(nonlin_df, train_df)
# test_df = anti_join(nonlin_df, train_df, by = "id")

ggplot(train_df, aes(x = x, y = y)) + 
  geom_point() + 
  geom_point(data = test_df, color = "red")
```

`mgcv::gam()`: Generalized additive models
* additive models e.g. polynomial models

Build three models with different compolexity (non-nested)
```{r}
linear_mod = lm(y ~ x, data = train_df)
smooth_mod = mgcv::gam(y ~ s(x), data = train_df)
wiggly_mod = mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)
```

`add_predictions(data, model)`: Add predictions to a data frame
`add_residuals(data, model)`: Add residuals to a data frame

```{r}
train_df %>% 
  add_predictions(linear_mod) %>% # add a column "pred"
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

`modelr::gather_predictions`: add predictions for several models to a data frame and then “pivoting” so the result is a tidy, “long” dataset that’s easily plottable. (add two columns: "model" and "pred")

```{r}
train_df %>% 
  gather_predictions(linear_mod, smooth_mod, wiggly_mod) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)
```

`rmse(model, data)` data used to get the predictions and then calculate rmse
```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```

## CV using `modelr`

```{r}
cv_df = 
  crossv_mc(nonlin_df, 100) # 100: number of splits => train(list), test(list), id

cv_df %>% pull(train) %>% .[[1]] %>% as_tibble # get the training set from the first split (id,x,y)
```


```{r}
cv_df = 
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) %>% # convert resample to tibble
  mutate(
    linear_mod  = map(.x = train, ~lm(y ~ x, data = .x)),
    smooth_mod  = map(.x = train, ~mgcv::gam(y ~ s(x), data = .x)),
    wiggly_mod  = map(.x = train, ~gam(y ~ s(x, k = 30), sp = 10e-6, data = .x))) %>% 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, ~rmse(model = .x, data = .y)),
    rmse_smooth = map2_dbl(smooth_mod, test, ~rmse(model = .x, data = .y)),
    rmse_wiggly = map2_dbl(wiggly_mod, test, ~rmse(model = .x, data = .y)))
```

