---
title: "linear_model"
author: "Jiayi Shi"
date: "2022-11-10"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(broom)

set.seed(1)
```

lm(formula, data): linear model
glm(): generalized linear model

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) %>% 
  filter(borough != "Staten Island") %>% 
  select(price, stars, borough, neighborhood, room_type) %>% 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type)
      )
```

fit model:

outcome ~ 1: intercept-only model.

outcome ~ 0 + ...: model with no intercept.

outcome ~ .: model using all available predictors.

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

categorical (factor) covariates: set reference for factor levels

```{r}
fit = 
  nyc_airbnb %>% 
  mutate( # change reference category
    borough = fct_infreq(borough) # set most common category as reference (most common category set as 1 by R - reference)
    ) %>% 
  lm(price ~ stars + borough, data = .)
```

```{r}
summary(fit)
summary(fit)$coef
coef(fit)
fitted.values(fit)
```

```{r}
fit %>% 
  broom::glance()
```

```{r}
fit %>% 
  broom::tidy() %>% 
  mutate(term = str_replace(term, "borough", "borough: ")) %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)
```

## Diagnostics

```{r}
modelr::add_residuals(nyc_airbnb, fit) %>%
  ggplot(aes(x = stars, y = resid)) +
  geom_point()

nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = borough, y = resid)) + 
  geom_violin() +
  ylim(-200, 250)

modelr::add_predictions(nyc_airbnb, fit)
```

## Hypothesis testing

```{r}
fit_null = lm(price ~ stars, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough, data = nyc_airbnb)

anova(fit_null, fit_alt) %>% # ANOVA works for nested models only
  broom::tidy() # small p-value, should include borough variable
```

## Room type by borough

Interactions:

```{r}
nyc_airbnb %>% 
  lm(price ~ stars + room_type * borough, data = .) %>% 
  broom::tidy() 
```

Fit borough-specific models associating price with rating and room type:

```{r}
nyc_airbnb %>% 
nest(df = -borough) %>% # nest within boroughs / nest everything except brough
mutate(
  models = map(.x = df, ~lm(price ~ stars + room_type, data = .x)),
  results = map(models, broom::tidy)) %>% 
select(-df, -models) %>% 
unnest(results) %>% 
select(borough, term, estimate) %>% 
mutate(term = fct_inorder(term)) %>% 
pivot_wider(
  names_from = term, values_from = estimate) %>% 
knitr::kable(digits = 3)
```

```{r}
manhattan_airbnb =
  nyc_airbnb %>% 
  filter(borough == "Manhattan")

manhattan_nest_lm_res =
  manhattan_airbnb %>% 
  nest(data = -neighborhood) %>% 
  mutate(
    models = map(data, ~lm(price ~ stars + room_type, data = .x)),
    results = map(models, broom::tidy)) %>% 
  select(-data, -models) %>% 
  unnest(results)

manhattan_nest_lm_res %>% 
  filter(str_detect(term, "room_type")) %>% 
  ggplot(aes(x = neighborhood, y = estimate)) + 
  geom_point() + 
  facet_wrap(~term) + 
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```

